---
title: "Statistics: Team Assignment"
output: html_notebook
author: "Maximilian Grotz"
---

# Data Preparation 

Change import of sample file and insert  $\alpha$ (alpha). 

```{r warning = FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)

sample <- read_excel("Sample3.xlsx", 
    col_types = c("numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "skip", "skip", "skip"))

alpha = 0.07
```

# Question 1

Determine for each of the variables EMPLOYEE, PROFIT, DCOSTS and GROWTH whether it is a qualitative or quantitative variable. Furthermore, for each qualitative variable indicate whether it is nominal or ordinal and for each quantitative variable indicate whether it is discrete or continuous.

| Variable | Type         | Specific   |
| ---------| -------------| -----      |
| Employee | Quantitative | Discrete   |
| Profit   | Quantitative | Continuous |
| dCost    | Qualitative  | Nominal    |
| Growth   | Qualitative  | Ordinal    |


# Question 2

Compute the sample median, the sample mean, the sample variance and the sample standard deviation for the variable EMPLOYEE.

```{r}
median_EMP <- median(sample$EMPLOYEE)
mean_EMP <- mean(sample$EMPLOYEE)
svar_EMP <- var(sample$EMPLOYEE)
ssd_EMP <- sd(sample$EMPLOYEE)

# Calculating the sample variance "by hand"

variance_sample <- sample %>%
  mutate(dev_squared = (EMPLOYEE - mean_EMP)^2)
manual_svar <- sum(variance_sample$dev_squared) / (nrow(variance_sample) - 1)


cat(paste("The median is", round(median_EMP, 4),
          "\nThe mean is", round(mean_EMP, 4),
          "\nThe sample variance is", round(svar_EMP, 4),
          "\nThe (manual) sample variance is", round(manual_svar, 4), 
          "\nThe sample standard deviation is", round(ssd_EMP, 4), 
          "\nThe (manual) sample standard deviation is", round(manual_svar^0.5,4)))
```


# Question 3

Compute the sample median and mode for the variable GROWTH (Use the numerical values assigned to the categories; 1 = low, 2 = medium, 3 = high.

```{r}
median_growth <- median(sample$GROWTH)

mode <- function(x){which.max(tabulate(x))}
mode_growth <- mode(sample$GROWTH)


cat(paste("The median is", median_growth, "\nThe mode is", mode_growth))
```



# Question 4

Determine the relative frequencies of “Technology”, “Finance”, “Healthcare” and “Other” for the variable SECTOR.

```{r}
relfreq_SECTOR <- table(sample$SECTOR) / length(sample$SECTOR)

index_relfreq <- 1

for (type in c("Technology", "Finance", "Healthcare", "Other")) {
  print(paste("The relative frequency of", type, "is", round(relfreq_SECTOR[index_relfreq], 4)))
  index_relfreq <- index_relfreq +1
}
```


# Question 5

Determine the sample mean and sample standard deviation of the variable COMPETITORS. Compute also the sample proportion of companies that have less than 5 competitors

```{r}
mean_COMP <- mean(sample$COMPETITORS)
ssd_COMP <- sd(sample$COMPETITORS)

prop_less5 <- sum(sample$COMPETITORS < 5) / length(sample$COMPETITORS)


cat(paste("The sample mean of COMPETITORS is", round(mean_COMP, 4), 
          "\nThe sample standard deviation of COMPETITORS is", round(ssd_COMP, 4), 
          "\nThe sample proportion of companies that have less than 5 competitors is", round(prop_less5, 4)))
```



# Question 6

1. Compute for the variable PROFIT the intervals around the sample mean with 1.25, 1.5, 1.75 and 2 sample standard deviations. In other words, compute ($\bar{x}$ − 1.25s, $\bar{x}$ + 1.25s), ($\bar{x}$ − 1.5s, $\bar{x}$ + 1.5s), ($\bar{x}$ − 1.75s, $\bar{x}$ + 1.75s) and ($\bar{x}$ − 2s, $\bar{x}$ + 2s). Report the lower and upper bounds of these intervals. 

2. Compute the minimum proportion of observations included in these intervals according to Chebyshev’s rule.

3. Compute the exact proportions of observations included in these intervals according to the actual data.

```{r}
mean_profit <- mean(sample$PROFIT)
sd_profit <- sd(sample$PROFIT)

sd_multiples <- c(1.25, 1.5, 1.75, 2)
q6 <- data.frame(sd_multiples) %>%
  mutate(lowbound = mean_profit - sd_profit * sd_multiples,
         highbound = mean_profit + sd_profit * sd_multiples, 
         minCheb = 1 - 1 / sd_multiples^2)

for (i in 1:4){
  q6$prop[i] <- sum((sample$PROFIT > q6$lowbound[i]) & (sample$PROFIT < q6$highbound[i])) / nrow(sample)
}
round(q6,4)
```



# Question 7 

1. Compute the z-scores for the variable PROFIT. Report the lowest and the highest z-score of the variable PROFIT in the sample and the value of that observation. Determine whether the value with the lowest z-score and the highest z-score are outlier values or not by using the 3$\sigma$ outlier detection method.

2. Compute the sample interquartile range (IQR) for the variable PROFIT. Determine the 
number of outliers using quartiles and IQR.

```{r}
findz_PROFIT <- function(x){
  (x - mean_profit) / sd_profit
}

lowest_z <- findz_PROFIT(min(sample$PROFIT))
highest_z <- findz_PROFIT(max(sample$PROFIT))

cat("The lowest z-score is", lowest_z, 
          "\nThe highest z-score is", highest_z, 
          "\nIt is", abs(lowest_z) > 3, "that the lowest z-score is an outlier value. ", 
          "\nIt is", abs(highest_z) > 3, "that the highest z-score is an outlier value.")


iqr_profit <- IQR(sample$PROFIT)
Q1_profit <- quantile(sample$PROFIT, 0.25)
Q3_profit <- quantile(sample$PROFIT, .75)

number_outliers_profit <- sum(sample$PROFIT < Q1_profit-1.5*iqr_profit) + sum(sample$PROFIT > Q3_profit+1.5*iqr_profit)

cat("The number of outliers is", number_outliers_profit)
```



# Question 8 

a) Determine (1-$\alpha$) × 100% -confidence intervals for the population means µ5 and µ6 of the 
variables 5 and 6 (PROFIT and COMPETITORS, respectively). Report the lower and upper 
bounds of the intervals. 
b) Use these confidence intervals to provide the statistical conclusions of the following three 
testing problems:
    1. $H_0$: $\mu_5 \le 0$ against $H_1$: $\mu_5 > 0$
    2. $H_0$: $\mu_6 \ge 6$ against $H_1$: $\mu_6 < 6$

```{r}
t_alpha2_300 <- qt(1 - alpha/ 2, 300)

lowbound_profit <- mean_profit - t_alpha2_300 * sd_profit / nrow(sample)^0.5
highbound_profit <- mean_profit + t_alpha2_300 * sd_profit / nrow(sample)^0.5

lowbound_COMP <- mean_COMP - t_alpha2_300 * ssd_COMP / nrow(sample)^0.5
highbound_COMP <- mean_COMP + t_alpha2_300 * ssd_COMP / nrow(sample)^0.5


H0_profit <- ifelse(highbound_profit < 0, "Do not reject H0", "Reject H0")
H0_COMP <- ifelse(lowbound_COMP > 6, "Do not reject H0", "Reject H0")

cat("The lower bound of the CI for profit is", round(lowbound_profit, 4), "and the higher bound is", round(highbound_profit, 4), 
    "\nThe lower bound of the CI for competitors is", round(lowbound_COMP, 4), "and the higher bound is", round(highbound_COMP, 4), 
    "\n\nPROFIT:\n", H0_profit, 
    "\n\nCOMPETITORS:\n", H0_COMP)

```


# Question 9

Consider the following two testing problems:

1. $H_0$: $p_7 \le 0.70$ against $H_1$: $p_7 > 0.7$
2. $H_0$: $p_8 = 0.5$ against $H_1$: $p_8 \neq 0.5$

where $p_7$ and $p_8$ refer to the population proportions of the companies that reduced fixed costs in 
2020, and hired employees in 2020, respectively. Report the corresponding value of test statistics and p-values, and provide the statistical conclusions

```{r}
mean_DCOSTS <- mean(sample$DCOSTS)
TestStat_DCOSTS <- (mean_DCOSTS - 0.7) / ((0.7 * (1 - 0.7)/nrow(sample))^0.5)
pval_DCOSTS <- pnorm(-TestStat_DCOSTS, 0, 1)

mean_DNEWHIRES <- mean(sample$DNEWHIRES)
TestStat_DNEWHIRES <- (mean_DNEWHIRES - 0.5) / ((0.5 * (1 - 0.5)/nrow(sample))^0.5)
pval_DNEWHIRES <- 2 * ifelse(TestStat_DNEWHIRES >0, pnorm(TestStat_DNEWHIRES, 0, 1, lower.tail = FALSE), pnorm(TestStat_DNEWHIRES, 0, 1, lower.tail = TRUE))

H0_DCOSTS <- ifelse(pval_DCOSTS > alpha, "Do not reject H0", "Reject H0")
H0_DNEWHIRES <- ifelse((pval_DNEWHIRES > alpha), "Do not reject H0", "Reject H0")

cat("DCOSTS (Companies that reduced fixed costs in 2020)", 
    "\nTest statistic:", TestStat_DCOSTS, 
    "\np-value:", pval_DCOSTS, 
    "\nStatistical conclusion:", H0_DCOSTS,
    "\n\nDNEWHIRES (Companies that hired employees in 2020)",
    "\nTest statistic:", TestStat_DNEWHIRES, 
    "\np-value:", pval_DNEWHIRES, 
    "\nStatistical conclusion:", H0_DNEWHIRES)
```



# Question 10

1. $H_0$: $\mu_2 \le 11$ against $H_1$: $\mu_2 > 11$
2. $H_0$: $\mu_6 \ge 100$ against $H_1$: $\mu_6 < 100$

where, $µ_2$ and $µ_4$ refer to the population means of the variables 2 and 4 (AGE and EMPLOYEE, respectively). Report the critical values determining the rejection regions and indicate the rejection regions. Report the value of test statistics and provide the statistical conclusions.

```{r}
mean_AGE <- mean(sample$AGE)
sd_AGE <- sd(sample$AGE)
t_alpha_300_AGE <- qt(1-alpha, 300)
TestStat_AGE <- (mean_AGE - 11)/(sd_AGE / nrow(sample)^0.5)

H0_AGE <- ifelse(t_alpha_300_AGE > TestStat_AGE, "Do not reject H0", "Reject H0")

cat("Age:", 
    "\nThe critical value is", t_alpha_300_AGE,
    "\nThe rejection region is to the right side of the critical value.",
    "\nThe value of the test statistic is", TestStat_AGE, 
    "\nThe statistical conclusion is:", H0_AGE, 
    "\n\n")

t_alpha_300_EMP <- qt(alpha, 300)
TestStat_EMP <- (mean_EMP - 100)/(ssd_EMP / nrow(sample)^0.5)

H0_EMP <- ifelse(t_alpha_300_EMP < TestStat_EMP, "Do not reject H0", "Reject H0")

cat("Employment:", 
    "\nThe critical value is", t_alpha_300_EMP,
    "\nThe rejection region is to the left side of the critical value.",
    "\nThe value of the test statistic is", TestStat_EMP, 
    "\nThe statistical conclusion is:", H0_EMP)
```

